\documentclass[House_Rules.tex]{subfiles}
\begin{document}

\chapter{Playtest Chr. Options}

You can choose to use this material when creating a character. However, balancing changes may be required. 

After sufficient playtesting, features here may be added to the core character options.




\section{Sources}

\begin{DndTable}[]{Xl}
    \textbf{Sourcebook} & \textbf{Sections} \\
    Explorer's Guide to Wildemount (EGW) & Backgrounds, spells, subclasses \\
\end{DndTable}

\begin{DndTable}[header=Unearthed Arcana]{lX}
    \textbf{Document} & \textbf{Sections} \\
    Class Feature Variants (UACV) & Ranger only \\
\end{DndTable}




\section{Races}

\subsection{Dragonborn}
\textit{Amends Breath Weapon (PH, 34).}

\subparagraph{Breath Weapon} Your breath weapon uses a bonus action. Its damage is 2d4 at 1st level, 3d4 at 6th level, 4d4 at 11th level, and 5d4 at 16th level.

\subsection{Human}

You gain these traits in addition to your base traits (PH, 31). You should not use variant humans.

\subparagraph{Versatile} You gain proficiency in one skill, one tool, one simple weapon, and one martial weapon of your choice.
\subparagraph{Talented} You can double your proficiency bonus in one skill from your class list in which you are proficient.




\section{Classes}

\subsection{Barbarian}

\subsubsection{Frenzy}
\textit{Replaces Berserker's Frenzy (PH, 49).}

Starting when you choose this path at 3rd level, you can go into a frenzy when you rage. If you do so, for the duration of your rage you can attack one additional time whenever you take the Attack action on your turn. When your rage ends, you suffer one level of exhaustion.

\subsection{Monk}
\subsubsection{Stunning Strike}
\textit{Replaces Stunning Strike (PH, 79).}

Starting at 5th level, you have one use of this feature and can expend it on your turn to make all your unarmed strikes stunning strikes until the end of that turn. If a stunning strike hits a creature, it must succeed on a Constitution saving throw or be stunned until the end of your next turn. When you finish a short or long rest, you regain all expended uses. You gain a second use of this feature at 11th level and a third at 17th level. 

\subsection{Ranger}
You may use a Variant Ranger instead of the other variations.

A Variant Ranger is an Original Ranger amended with these features from Unearthed Arcana (UACV, 7-9):

\begin{itemize}
    \item Deft Explorer
    \item Favored Foe
    \item Fighting Style Options
    \item Ranger Spells
    \item Spellcasting Focus
    \item Primal Awareness
    \item Fade Away
    \item Ranger Companion Options
\end{itemize}

In particular, you should not use the Spell Versatility feature (UACV, 8).

You can also use the new Fighting Styles and the Martial Versatility feature (UACV, 12), but only through Variant Ranger class features.

\subsection{Warlock}

\subsubsection{Hexblade}
The Hex Warrior feature (XGE, 55) is removed from the Hexblade Patron and added to the Pact of the Blade (PH, 107).

You can multiclass with such a Hexblade Warlock.



\end{document}