\documentclass[House_Rules.tex]{subfiles}
\begin{document}

\chapter{Homebrew}

\section{Items}
\DndItemHeader{Bag of Many Goats}{Wondrous item, Major, Rare}
This ordinary bag, made from leather, appears empty. Reaching inside the bag, however, reveals the presence of a small curved horn.

As an action, you can expend a number of charges to pull the horn from the bag and throw it up to 20 feet. When the horn lands, it transforms into a magical goat, which vanishes at dawn or when it is reduced to 0 hit points.

The statistics of the goat are determined by the number of charges expended:

\begin{DndTable}[]{ccc}
    \textbf{Charges} & \textbf{Goat Size} & \textbf{Goat Speed} \\
    1 & Tiny & 15 feet \\
    1 & Small & 30 feet \\
    2 & Medium & 40 feet \\
    4 & Large & 40 feet
\end{DndTable}

The bag has a maximum of 4 charges, and regains 1d4 charges at dawn. If you try to expend more charges than there are charges remaining, your action is wasted.

The goats have normal vision. The goat's statistics other than speed and vision are the same as an \textbf{animated object} (PH, 213) of the same size.

The goats take their turn immediately after you. As a bonus action, you can command one goat verbally. You decide what action the goat will take and where it will move in its next turn. If you issue no command, the creature does nothing.

\DndItemHeader{Potion of Restfulness}{Potion, Minor, Uncommon}
As an action, a creature can drink this potion, immediately falling unconscious for 10 minutes. The creature wakes up if it takes any damage, or if another creature uses an action to shake it awake. If the creature remains unconscious for the entire duration, it then awakens and gains the benefit of a short rest.


\end{document}